Preprocess

### 2023-11-19 src/metrics.Rmd
Calculated metrics
Working pccc loop metrics

@@

@@

```{r}
calc_pccc <- function(actual, pred, k = 1, N = NULL) {

@@

calc_csmf_acc <- function(actual, pred) {

@@

    # Calc csmf acc
    out <- 1 - (sum(abs(csmf_true - csmf_pred)) / csmf_max_error)
    return(out)
}
```

@@

@@

```{r}
df <- read_csv("../data/healsl_rd1to2_cod_v1.csv")
```

```{r, echo=FALSE}

# Display num of cases
cat("Number of Cases: ", nrow(df), "\n")

# Get actual and pred
actual <- df$physician_cghr10
pred <- df %>% select(ends_with("_cghr10"), -physician_cghr10)

# Physician info
cat(paste0(
    "Physician Coded Cases (",
    length(unique(na.omit(actual))), " CODs): ",
    length(actual) - sum(is.na(actual) | is.null(actual)),
    "\n"
))

# Models info
for (mcol in colnames(pred)) {
    cat(paste0(
        toupper(str_remove(mcol, "_cghr10")), " Predicted Cases (", 
        length(unique(na.omit(pred[[mcol]]))), " CODs): ",
        length(pred[[mcol]]) - sum(is.na(pred[[mcol]]) | is.null(pred[[mcol]])),
        "\n"
    ))
}

print(head(df))
```

@@

```{r}
df <- df %>%
    filter(!is.na(physician_cghr10) & !is.null(physician_cghr10))
```

```{r, echo=FALSE}
cat("Number of Cases for Metrics: ", nrow(df))
print(head(df))
```

@@

Separate physician (actual) and model (predicted) CODs.

```{r}
actual <- df$physician_cghr10
pred <- df %>%
    select(ends_with("_cghr10"), -physician_cghr10) %>%
    rename_all(~toupper(str_replace(., "_cghr10", "")))
```

```{r, echo=FALSE}
cat(paste0(
    "Physician Coded Cases (",
    length(unique(na.omit(actual))), " CODs): ",
    length(actual) - sum(is.na(actual) | is.null(actual)),
    "\n"
))
```

```{r, echo=FALSE}
for (mcol in colnames(pred)) {
    cat(paste0(
        mcol, " Predicted Cases (", 
        length(unique(na.omit(pred[[mcol]]))), " CODs): ",
        length(pred[[mcol]]) - sum(is.na(pred[[mcol]]) | is.null(pred[[mcol]])),
        "\n"
    ))
}
```

@@

@@

Calculate PCCC to evaluate indivudal performance for each model.

```{r}
pccc <- pred %>%
    pivot_longer(
        everything(),
        names_to = "Model",
        values_to = "prediction"
    ) %>%
    group_by(Model) %>%
    summarise(
        PCCC = calc_pccc(actual, prediction, k = str_count(unique(Model), "_") + 1)
    ) %>%
    mutate(Model = str_replace_all(Model, "_", " & ")) %>%
    arrange(desc(PCCC))
pccc
```

@@

Calculate CSMF Accuracy to evaluate population performance for each model.

```{r}
csmf_acc <- pred %>%
    pivot_longer(
        everything(),
        names_to = "Model",
        values_to = "prediction"
    ) %>%
    group_by(Model) %>%
    summarise(
        "CSMF Accuracy" = calc_csmf_acc(actual, prediction)
    ) %>%
    mutate(Model = str_replace_all(Model, "_", " & ")) %>%
    arrange(desc(`CSMF Accuracy`))
csmf_acc
```

>>>>

```{r, echo=FALSE, eval=FALSE}
install.packages("tidyverse")
```

@@

## Settings

```{r}

# Name formatting vars
models_format <- c(
    "_cghr10" = "",
    "_" = " & ",
    "chatgpt3" = "ChatGPT-3",
    "chatgpt4" = "ChatGPT-4",
    "insilicova" = "InSilicoVA",
    "interva5" = "InterVA-5"
)
```

@@

```{r}

# Calc PCCC
calc_pccc <- function(actual, pred, k = 1, N = NULL) {

@@

# Calc CSMF accuracy
calc_csmf_acc <- function(actual, pred) {

@@

    # Calc csmf acc
    out <- 1 - (sum(abs(csmf_true - csmf_pred)) / csmf_max_error)
    return(out)
}

# Bulk calc per model
calc_per_model <- function(actual, pred, func, name = "Metric", ...) {

    # Prep data in long format grouped by model
    out <- pred %>%
        pivot_longer( # to long format
            everything(),
            names_to = "Model",
            values_to = "prediction"
        ) %>% group_by(Model)

    # Calc k if pccc as it is diff for model combos
    if (identical(func, calc_pccc)) {
        out <- out %>%
            summarise(
                !!name := func(
                    actual,
                    prediction,
                    k = str_count(unique(Model), "_") + 1,
                    ...
                )
            )
    } else { # otherwise apply func
        out <- out %>%
            summarise(
                !!name := func(
                    actual,
                    prediction,
                    ...
                )
            )
    }

    # Format display and return
    out <- out %>%
        mutate(Model = str_replace_all(Model, models_format)) %>%
        arrange(across({{name}}, desc))
    return(out)
}
```

@@

### Raw Data

@@

```{r}
raw_df <- read_csv("../data/healsl_rd1to2_cod_v1.csv")
```

```{r, echo=FALSE}
cat("Cases: ", nrow(raw_df), "\n")
print(head(raw_df))
```

### Physician Cases

@@

```{r}
df <- raw_df %>%
    filter(!is.na(physician_cghr10) & !is.null(physician_cghr10))
```

```{r, echo=FALSE}
cat("Physician Coded Cases: ", nrow(df))
print(head(df))
```

### Combined Models

@@

### Prepared Data

Display case counts for raw data and prepared data after combining models and filtering for physician coded cases.

```{r}

cat("\nRaw Data\n--------\n\n")

# Physician info
raw_physicians <- raw_df$physician_cghr10
cat(paste0(
    "Cases (",
    length(unique(na.omit(raw_physicians))), " CODs): ",
    length(raw_physicians),
    "\n"
))

# Models cases
raw_models <- raw_df %>% select(ends_with("_cghr10"), -physician_cghr10)
for (mcol in colnames(raw_models)) {
    m <- raw_models[[mcol]]
    cat(paste0(
        str_replace_all(mcol, models_format), " Predicted Cases (", 
        length(unique(na.omit(m))), " CODs): ",
        length(m) - sum(is.na(m) | is.null(m)),
        "\n"
    ))
}

cat("\nPrepared Data\n--------------\n")

# Physician cases
physicians <- df$physician_cghr10
cat(paste0(
    "Physician Coded Cases (",
    length(unique(na.omit(physicians))), " CODs): ",
    length(physicians) - sum(is.na(physicians) | is.null(physicians)),
    "\n"
))

# Models cases
models <- df %>% select(ends_with("_cghr10"), -physician_cghr10)
for (mcol in colnames(models)) {
    m <- models[[mcol]]
    cat(paste0(
        str_replace_all(mcol, models_format), " Predicted Cases (", 
        length(unique(na.omit(m))), " CODs): ",
        length(m) - sum(is.na(m) | is.null(m)),
        "\n"
    ))
}
```

@@

Calculate metrics from model outputs compared to physician codes.

@@

Calculate Partial Chance Corrected Concordance (PCCC) to evaluate indivudal performance for each model.

```{r}
pccc <- list()

# Calc overall pccc
pccc$all <- calc_per_model(
    actual = df %>% pull(physician_cghr10),
    pred = df %>% select(ends_with("_cghr10"), -physician_cghr10),
    func = calc_pccc
) %>% rename(
    PCCC = Metric
)

# Calc pcc for each age group
for (age_group in c("adult", "child", "neo")) {

    # Filter data for age
    df_age <- df %>% filter(age == age_group)
    metric_name = paste0("PCCC ", str_to_title(age_group))

    # Calc metric for age
    pccc[[age_group]] <- calc_per_model(
        actual = df_age %>%
            pull(physician_cghr10),
        pred = df_age %>%
            select(ends_with("_cghr10"), -physician_cghr10),
        func = calc_pccc
    ) %>% rename(
         {{ metric_name }} := Metric
    )
}

# Combine all pccc metrics
pccc <- reduce(
    pccc,
    function(x, y) left_join(x, y, by = "Model")
) %>%
    arrange(desc(PCCC))
pccc
```

@@

Calculate Cause Specific Mortality Fraction (CSMF) Accuracy to evaluate population performance for each model.

```{r}
csmf_acc <- list()
```

#### Overall CSMF Accuracy

```{r}
csmf_acc <- calc_per_model(
    physicians,
    models,
    calc_csmf_acc,
    "CSMF Accuracy"
)
csmf_acc
```

#### Adult Accuracy

#### Child Accuracy

#### Neonate Accuracy

### 2023-11-19 src/preprocess.Rmd
Working pccc loop metrics

# Read data
df <- list(
    wva2016_icd10 = read_csv(
        "../data/wva2016_icd10_v1.csv",
        col_types = cols(wva2016_code = col_double())
    ),
    icd10_cghr10 = read_csv("../data/icd10_cghr10_v1.csv"),
    physician = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_adult_v1.csv",
            col_select = phy_adult_cols
        ) %>%
            mutate(round = 1, age = "adult") %>%
            rename(all_of(phy_adult_rename)),
        read_csv(
            "../tmp/healsl_rd2_adult_v1.csv",
            col_select = phy_adult_cols
        ) %>%
            mutate(round = 2, age = "adult") %>%
            rename(all_of(phy_adult_rename)),
        read_csv(
            "../tmp/healsl_rd1_child_v1.csv",
            col_select = phy_cols
        ) %>% mutate(round = 1, age = "child"),
        read_csv(
            "../tmp/healsl_rd2_child_v1.csv",
            col_select = phy_cols
        ) %>% mutate(round = 2, age = "child"),
        read_csv(
            "../tmp/healsl_rd1_neo_v1.csv",
            col_select = phy_cols
        ) %>% mutate(round = 1, age = "neo"),
        read_csv(
            "../tmp/healsl_rd2_neo_v1.csv",
            col_select = phy_cols
        ) %>% mutate(round = 2, age = "neo")
    ),
    model_chatgpt3 = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_rapid_chatgpt3_v1.csv",
            col_select = c("rowid", "round", "api_chat_icd")
        ),
        read_csv(
            "../tmp/healsl_rd2_rapid_chatgpt3_v1.csv",
            col_select = c("rowid", "round", "api_chat_icd")
        )
    ) %>% rename(chatgpt3_icd10 = api_chat_icd),
    model_chatgpt4 = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_rapid_chatgpt4_v1.csv",
            col_select = c("rowid", "round", "api_chat_icd")
        ),
        read_csv(
            "../tmp/healsl_rd2_rapid_chatgpt4_v1.csv",
            col_select = c("rowid", "round", "api_chat_icd")
        )
    ) %>% rename(chatgpt4_icd10 = api_chat_icd),
    model_insilicova = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_rapid_insilicova_v1.csv",
            col_select = c("rowid", "round", "cause1_wva")
        ),
        read_csv(
            "../tmp/healsl_rd2_rapid_insilicova_v1.csv",
            col_select = c("rowid", "round", "cause1_wva")
        )
    ) %>% rename(insilicova_wva2016 = cause1_wva),
    model_interva5 = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_rapid_interva5_v1.csv",
            col_select = c("rowid", "round", "cause1_wva")
        ),
        read_csv(
            "../tmp/healsl_rd2_rapid_interva5_v1.csv",
            col_select = c("rowid", "round", "cause1_wva")
        )
    ) %>% rename(interva5_wva2016 = cause1_wva)
)
```

@@

## Method

Create column for final physician ICD-10 codes based on the following rule:

@@

```{r}
df$physician <- df$physician %>%
    mutate(physician_icd10 = coalesce(
        adj_icd,
        p1_recon_icd,
        p1_icd
    ))
```

@@

>>>>

# Read data
df <- list(
    wva2016_icd10 = read_csv(
        "../data/wva2016_icd10_v1.csv",
        col_types = cols(wva2016_code = col_double())
    ),
    icd10_cghr10 = read_csv("../data/icd10_cghr10_v1.csv"),
    physician = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_adult_v1.csv",
            col_select = all_of(phy_adult_cols)
        ) %>%
            mutate(round = 1, age = "adult") %>%
            rename(all_of(phy_adult_rename)),
        read_csv(
            "../tmp/healsl_rd2_adult_v1.csv",
            col_select = all_of(phy_adult_cols)
        ) %>%
            mutate(round = 2, age = "adult") %>%
            rename(all_of(phy_adult_rename)),
        read_csv(
            "../tmp/healsl_rd1_child_v1.csv",
            col_select = all_of(phy_cols)
        ) %>% mutate(round = 1, age = "child"),
        read_csv(
            "../tmp/healsl_rd2_child_v1.csv",
            col_select = all_of(phy_cols)
        ) %>% mutate(round = 2, age = "child"),
        read_csv(
            "../tmp/healsl_rd1_neo_v1.csv",
            col_select = all_of(phy_cols)
        ) %>% mutate(round = 1, age = "neo"),
        read_csv(
            "../tmp/healsl_rd2_neo_v1.csv",
            col_select = all_of(phy_cols)
        ) %>% mutate(round = 2, age = "neo")
    ),
    model_chatgpt3 = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_rapid_chatgpt3_v1.csv",
            col_select = c("rowid", "round", "api_chat_icd")
        ),
        read_csv(
            "../tmp/healsl_rd2_rapid_chatgpt3_v1.csv",
            col_select = c("rowid", "round", "api_chat_icd")
        )
    ) %>% rename(chatgpt3_icd10 = api_chat_icd),
    model_chatgpt4 = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_rapid_chatgpt4_v1.csv",
            col_select = c("rowid", "round", "api_chat_icd")
        ),
        read_csv(
            "../tmp/healsl_rd2_rapid_chatgpt4_v1.csv",
            col_select = c("rowid", "round", "api_chat_icd")
        )
    ) %>% rename(chatgpt4_icd10 = api_chat_icd),
    model_insilicova = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_rapid_insilicova_v1.csv",
            col_select = c("rowid", "round", "cause1_wva")
        ),
        read_csv(
            "../tmp/healsl_rd2_rapid_insilicova_v1.csv",
            col_select = c("rowid", "round", "cause1_wva")
        )
    ) %>% rename(insilicova_wva2016 = cause1_wva),
    model_interva5 = bind_rows(
        read_csv(
            "../tmp/healsl_rd1_rapid_interva5_v1.csv",
            col_select = c("rowid", "round", "cause1_wva")
        ),
        read_csv(
            "../tmp/healsl_rd2_rapid_interva5_v1.csv",
            col_select = c("rowid", "round", "cause1_wva")
        )
    ) %>% rename(interva5_wva2016 = cause1_wva)
)
```

@@

## Methods

### Final ICD

Create one column for final physician ICD-10 codes and one column each for physician agreed, reconciled, and adjudicated records based on the following rule:

@@

```{r}
df$physician <- df$physician %>%
    mutate(
        physician_icd10 = coalesce(
            adj_icd,
            p1_recon_icd,
            p1_icd
        ),
        is_agreed = if_else(
            is.na(p1_recon_icd) & is.na(p2_recon_icd) & is.na(adj_icd),
            TRUE,
            FALSE
        ),
        is_recon = if_else(
            !is.na(p1_recon_icd) | !is.na(p2_recon_icd),
            TRUE,
            FALSE
        ),
        is_adj = if_else(
            !is.na(adj_icd),
            TRUE,
            FALSE
        )
    ) %>%
    select(
        rowid,
        round,
        age,
        p1_icd,
        p2_icd,
        p1_recon_icd,
        p2_recon_icd,
        adj_icd,
        is_agreed,
        is_recon,
        is_adj,
        physician_icd10
    )
head(df$physician)
```

### Combine Outputs

@@

### Convert to CGHR-10
